---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, eval = FALSE)
```

Now we want to port our Dollar Growth work to a Shiny application so that an end user is able to: 

1) choose tickers and portfolio weights
2) choose a start date
3) choose a rebalancing frequency
4) chart the growth of a dollar in the portfolio since the chosen start date

The final app can be seen here with full source code:

www.reproduciblefinance.com/shiny/portfolio-growth/

The input sidebar is identical to that of our app on returns distribtution We let the user choose 5 ticker symbols, 5 weights, a start date and rebalance period. The user then clicks 'submit' to fire up the reactives. Since it's identical, we're not going to review it here. 

We will, though, review the difference substantive flow that is used to get portfolio growth. We use an `eventReactive()`, pass in tickers, weights, start date and frequency, but instead of calculating portfolio returns, we calculate portfolio growth. It's not a copy paste from our first app, but it's a similar structure. Since we use `highcharter` in this app, we go with the `xts` world and use 
`Return.portfolio(asset_returns_xts, wealth.index = 1,...)` - this is the same method we used in our previous growth calculation.

```{r, eval = FALSE}
portfolio_growth_xts <- eventReactive(input$go, {
  
  symbols <- c(input$stock1, input$stock2, input$stock3, input$stock4, input$stock5)
  
  prices <- getSymbols(symbols, src = 'yahoo', from = input$date, 
             auto.assign = TRUE, warnings = FALSE) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>%
  `colnames<-`(symbols)

  w <- c(input$w1/100, input$w2/100, input$w3/100, input$w4/100, input$w5/100)

  prices_monthly <- to.monthly(prices, indexAt = "last", OHLC = FALSE)

  asset_returns_xts <- na.omit(Return.calculate(prices_monthly, method = "log"))

  portfolio_growth_xts <- 
    Return.portfolio(asset_returns_xts, 
                   wealth.index = 1, 
                   weights = w, 
                   rebalance_on = input$rebalance) %>% 
    `colnames<-`("growth")

})
```

At this point, portfolio growth is calculated and stored as `portfolio_growth_xts()` and our substantive work has been completed. Now we want to display the chart of the portfolio growth over time.  Outside of Shiny, this would be a simple passing of the xts object to `highcharter`. 

As with `ggplot`, Shiny uses a custom function for building reactive highcharter charts called `renderHighchart()`. Once we invoke that `renderHighchart()`, our code looks very similar to our previous visualization work as we use `hc_add_series(portfolio_growth_xts(), name = "Dollar Growth", color = "cornflowerblue")` to add our portfolio growth xts object to a chart.

```{r, eval = FALSE}
renderHighchart({
  
  highchart(type = "stock") %>%
  hc_title(text = "Growth of a Dollar") %>%
  hc_add_series(portfolio_growth_xts(), name = "Dollar Growth", color = "cornflowerblue") %>%
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE) 
})
```

Next, we use `ggplot()` to create the same visual as above but in the tidy world. The code flow is quite similar to how we would normally create a line chart, except we first need to convert our `xts` object to a tibble with `tk_tbl(preserve_index = TRUE, rename_index = "date")`.  Then we make our call to `ggplot(aes(x = date))` and add a geom with `geom_line(aes(y = growth), color = "cornflowerblue")`.

Note that the `renderPlot()` function is playing the same role as `renderHighchart()` above - it is alerting the Shiny app that a reactive plot is forthcoming after user inputs, instead of a static plot that is unchanging.

```{r, eval = FALSE}
renderPlot({
portfolio_growth_xts() %>% 
    tk_tbl(preserve_index = TRUE, rename_index = "date") %>% 
    ggplot(aes(x = date) +
    geom_line(aes(y = growth), color = "cornflowerblue") +
    ylab("dollars") + 
    ggtitle("Growth of Dollar over time")
})
```






