---
title: "portfolio dollar growth"
output:
  html_notebook: default
  pdf_document: default
---

```{r setup, message = FALSE, include = FALSE}

knitr::opts_chunk$set(message=FALSE, warning=FALSE)

library(tidyverse)
library(tidyquant)
library(timetk)
```

```{r, eval = TRUE, include = FALSE}
# The symbols vector holds our tickers. 
symbols <- c("SPY","EFA", "IJS", "EEM","AGG")

# The prices object will hold our raw price data throughout this book.
prices <- 
  getSymbols(symbols, src = 'yahoo', from = "2005-01-01", 
             auto.assign = TRUE, warnings = FALSE) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>%
  `colnames<-`(symbols)
# XTS method
prices_monthly <- to.monthly(prices, indexAt = "last", OHLC = FALSE)
asset_returns_xts <- na.omit(Return.calculate(prices_monthly, method = "log"))

# Tidyverse method, to long, tidy format
asset_returns_long <- 
  prices %>% 
  to.monthly(indexAt = "last", OHLC = FALSE) %>% 
  tk_tbl(preserve_index = TRUE, rename_index = "date") %>%
  gather(asset, returns, -date) %>% 
  group_by(asset) %>%  
  mutate(returns = (log(returns) - log(lag(returns))))

w <- c(0.25, 0.25, 0.20, 0.20, 0.10)

# Let's assign each asset a weight from our weights vector above.

w_1 <- w[1]
w_2 <- w[2]
w_3 <- w[3]
w_4 <- w[4]
w_5 <- w[5]

# And each asset has a return as well, stored in our 
# asset_returns_xts object.

asset1 <- asset_returns_xts[,1]
asset2 <- asset_returns_xts[,2]
asset3 <- asset_returns_xts[,3]
asset4 <- asset_returns_xts[,4]
asset5 <- asset_returns_xts[,5]

portfolio_returns_byhand <-   
  (w_1 * asset1) + 
  (w_2 * asset2) + 
  (w_3 * asset3) +
  (w_4 * asset4) + 
  (w_5 * asset5)

names(portfolio_returns_byhand) <- "returns"

portfolio_returns_xts_rebalanced_monthly <- 
  Return.portfolio(asset_returns_xts, weights = w, rebalance_on = "months") %>%
  `colnames<-`("returns")

portfolio_returns_tq_rebalanced_monthly <- 
  asset_returns_long %>%
  tq_portfolio(assets_col  = asset, 
               returns_col = returns,
               weights     = w,
               col_rename  = "returns",
               rebalance_on = "months")
```



We have covered the process of importing daily price data for 5 assets, converting to monthly log returns for those assets, and then converting to portfolio returns after assigning weights to those assets. Next we want to convert those portfolio returns to the growth of a dollar over time, so that each month's observation is not a monthly return but rather how a dollar invested in the porfolio in month 1 would have grown cumulatively.

There are a few motivations here. One is to make more intuitive, two is this is what people care about, three it's a nice visualization over time and will lay the groundwork for our first shiny app, and that lays the ground work for all of our shiny apps.


```{r}
portfolio_growth_byhand <- cumprod(1 + portfolio_returns_byhand$returns)
```

The XTS world two ways. 

First method is translate direct from asset returns to portfolio growth with the same `Return.portfolio()` function as we used to calculate portfolio returns. We pass in the weights vector as before, but we include the argument `wealth.index = 1`. This tells the function to calculate the growth of a dollar, as if our wealth started at $1 invested. 

```{r}
portfolio_growth_xts_skip_step <- 
  Return.portfolio(asset_returns_xts, 
                   wealth.index = 1, weights = w, rebalance_on = "months") %>%
  `colnames<-`("growth") 
```


The second method makes use of the portfolio returns object that we calculated earlier. We take the object `portfolio_returns_xts_rebalanced_monthly` and pass it to `Return.portfolio()`. We don't need to supply weights but we do again set `wealth.index = 1`. 

```{r}

portfolio_growth_xts_rebalanced_monthly <- 
  Return.portfolio(portfolio_returns_xts_rebalanced_monthly, 
                   wealth.index = 1) %>%
  `colnames<-`("growth") 

```

These two methods yield the same result for the growth of a dollar in our portfolio.  At first glance, the first method might seem to be better. It is more concise because we skip the step of converting asset returns to portfolio returns. But that skipping has a cost because we no longer have an object of monthly portfolio returns, and that means we cannot visualize those returns or calculate their standard deviations, skewness etc. 

This illustrates a subtle risk of relying on the highest level built in function to accomplish a task. That highest level function, which wraps so many other tasks, also elides intermediate data objects that might be useful to ourselves or others who see our end results but want to dig into our derivation of those end results.

A briefi look back to the idea of work flow and code flow. The two `xts` methods, as noted, give us the same result for the growth of a dollar. We can chart it, share it, use it in production.  

On to a tidyquant method for calculating growth of a dollar. Very similar to the `xts` method, we can go direct from asset returns to portfolio growth with the following code chunk. We use the `tq_portfolio()` function, pass it `asset_returns_long`, along with a weights vector and `wealth.index = 1`.

```{r}
portfolio_growth_tq_rebalanced_monthly <- 
  asset_returns_long %>%
  tq_portfolio(assets_col  = asset, 
               returns_col = returns,
               weights     = w,
               col_rename  = "growth",
               rebalance_on = "months", 
               wealth.index = 1) %>% 
  slice(-1)
  
```

That gives us consistent results with the `xts` method but again it skipped our important step of first finding monthly returns. Let's combine the tidyverse and tidyquant for a code flow still relies on those monthly returns. 

```{r}
portfolio_growth_tidy <- 
  portfolio_returns_tq_rebalanced_monthly %>% 
  mutate(growth = cumprod(1 + returns)) %>% 
  select(-returns) %>% 
  slice(-1)
```



Let's explore one more method using the `purrr` package from the tidyverse. `purrr` contains a lot of useful functions and the `map()` family is the most used, but another very useful function is `accumulate()`. As the name implies, this function allows us to accumulate calculations in a recursive way, meaning we can use prior calculations the same way we would in a for loop or with `cumprod()`. We will make use of this when we get to monte carlo simulations but here let's use `accumulate()` to get the growth of a dollar. 

```{r}
portfolio_growth_purrr <- 
  portfolio_returns_tq_rebalanced_monthly %>%  
  slice(-1) %>% 
  mutate(growth_1 = accumulate(1 + returns, `*`),
         growth_2 = accumulate(1 + returns, function(x, y) x * y)) %>% 
  select(date, growth_1, growth_2)
```

Note that we used `accumulate()` in two ways: `accumulate(1 + returns, `*`)` and 
`accumulate(1 + returns, function(x, y) x * y)` and thusly have introduced a new package to the toolkit but we did so as part of a workflow on dollar growth that also uses familiar packages and functions.  

This is a good time to discuss best practice for when we want to introduce a new function or package to our team's work. Most importantly, we want to explicitly flag it and explain it. Without that, our collaborators might miss the new function or package altogether, especially if it is an internal package. Next, when possible, use the new tool to accomplish a task that we also have accomplished in a traditional way. Here, we introduce `accumulate()` to calculate growth of a dollar, having already done so with `cumprod()` and built-ins from `PerformanceAnalytics` and `tidyquant`.  We have created a familiar environment in which to introduce a new tool and if in the future we use that tool in a more complex environment (as we will when running monte carlo simulations), our collaborators will have got a prior introduction.  That helps make our more complex work more understandable and reproducible.


Back to the substance, we have now calculated the growth of a dollar in several ways: using `cumprod()` and `dplyr`, using the built-in functions with `perfromanceAnalytics` and `tidyquant`, and using the `accumulate()` function from `purrr`. Let's confirm that we got consistent results.  

```{r}
final_comparison <- 
  merge.xts(portfolio_growth_byhand, 
            portfolio_growth_xts_skip_step, 
            portfolio_growth_xts_rebalanced_monthly) %>% 
  tk_tbl(preserve_index = TRUE, rename_index = "date") %>% 
  bind_cols(portfolio_growth_tq_rebalanced_monthly) %>% 
  select(-date1) %>% 
  bind_cols(portfolio_growth_tidy) %>% 
  select(-date1) %>% 
  bind_cols(portfolio_growth_purrr) %>% 
  select(-date1) %>%
  `colnames<-`(c("date", "by_hand", "xts1", "xts2", "tq", "tidy",  "purrr1", "purrr2"))


head(final_comparison)
tail(final_comparison)
```

7 routes to the same results for dollar growth! A lot of grinding but should any of our colleagues wish to reproduce, resuse or extend our results, a plethora of code paths is available to them. Why might those paths be important? On a team of 10 R/finance ninjas, there's a good chance that one coder will prefer the `xts` world, one will prefer tidyquant and/or one will prefer the tidyverse and hand-rolled functions. It's not the case that every document needs to include all these different paths but it's good practice to have one file in the team's library that can be used as a global reference point. If language analogies are appealing, we can think of this as a Rosetta Script or where to turn when someone wonders what's the equivalent of this analytical path using another set of packages. 

Olivia is 6 years old. 
Roxanne is 4 years old. 
Eloisa is 1 year old. 

Beyond the flexibility of different code paths, the various object structures lend themselves to different visualizations techniques, which is an reason that some coders might prefer different paths. If `highcharter` is your axe, `xts` can be very appealing indeed. Let's see why. 

To use `highcharter`, we first call `highchart(type = "stock")` and then supply our `portfolio_growth_xts_rebalanced_monthly` object to the `hc_add-series()` function - we don't need to worry about the date index - and presto.

```{r}
highchart(type = "stock") %>% 
  hc_add_series(portfolio_growth_xts_rebalanced_monthly) 
```

We can add more aesthetics as always. I want a title and add it with `hc_title(text = "Growth of a Dollar")` and the y-axis label to be on left-hand side with a `$` sign. We add that with  `hc_yAxis(title = list(text = "growth of dollar"), opposite = FALSE, labels = list(format = "${value}")`. For whatever reason I don't like that date scroll bar at the bottom so I will set `hc_scrollbar(enabled = FALSE)`.

```{r}

highchart(type = "stock") %>% 
  hc_title(text = "Growth of a Dollar") %>%
  hc_add_series(portfolio_growth_xts_rebalanced_monthly, 
                  name = "Portfolio", color = "cornflowerblue", lineWidth = 1) %>%
  hc_yAxis(title = list(text = "growth of dollar"),
           opposite = FALSE,
           labels = list(format = "${value}")) %>% 
  hc_add_theme(hc_theme_flat()) %>%
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE)
```


That's a nice and efficient chart.  An end user can quickly discern what would have happened to a dollar over time, how the financial crisis affected the portfolio and how things have recovered since.

We can also use `ggplot()` for time series plotting if we wish to stay in the tidy world. We use our `date` column as the x-axis and `growth_1` as the y-axis by calling `ggplot(aes(x = date, y = growth_1))`. We will add a `$` sign to the y-axis label with `scale_y_continuous(breaks = pretty_breaks(n = 10), labels = dollar)` which requires loading the `scales` package.

```{r}
library(scales)
portfolio_growth_purrr %>%
  ggplot(aes(x = date, y = growth_1)) + 
  geom_line(colour = "cornflowerblue") + 
  ylab("dollar growth") +
  xlab("month") +
  ggtitle("Growth of a Dollar") +
  scale_y_continuous(breaks = pretty_breaks(n = 10), labels = dollar) +
  scale_x_date(breaks = pretty_breaks(n = 10))
```


The end result with `ggplot` or `highcharter` does not vary a lot though highcharter does offer a bit more interactivity and the built-in date selection buttons. 



### On to Shiny

