---
title: "Sharpe Ratio Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r, include = FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
```

Today, we begin a project to build a Shiny application that allows a user to build a portfolio and calculate/visualize its Sharpe Ratio. 

The final app is viewable [here]()) but we'll spend the next 3 posts contructing that.

Briefly, the Sharpe Ratio is the mean of the excess portfolio returns above the risk-free rate, divided by the standard deviation of the excess monthly returns above the risk-free rate.  This is the formulation of the Sharpe Ratio as of 1994; if we wished to use the original formulation from 1966 the denominator would be the standard deviation of portfolio monthly returns. Learn more [here](http://web.stanford.edu/~wfsharpe/art/sr/sr.htm).

In other words, the Sharpe Ratio measures excess returns per unit of volatility, where we take the standard deviation to represent portfolio volatility. The Sharpe Ratio was brought to us by Bill Sharpe - arguably the most important economist for modern investment management as the creator of the Sharpe Ratio, CAPM and Financial Engines, a forerunner of today's robo-advisor movement.


The Sharpe Ratio eqation is as follows:   

$$Sharpe~Ratio={(\overline{R_{p}-R_{f}})}/\sigma_{excess}$$

The numerator is the mean excess return above the risk free rate and the numerator is the standard deviation of those excess returns.  In other words, it's a ratio of return to risk and so a higher Sharpe Ratio indicates a 'better' portfolio.  .



As for our project, we will proceed in 3 steps:

    1. Build a portfolio and calculate the Sortino Ratio using 3 methods (today's post)
    2. Visualize the Sortino Ratio using ggplot and highcharter (next week)
    3. Wrap to an interactive Shiny App (in two weeks)

When working with the Sharpe Ratio, we have two critical choices: how to construct the portfolio using assets and weights, and which risk free to use.  Our Shiny application at project's end will allow a user to make these choices and see how the Sortino Ratio changes.  For today, we will go with the following portfolio and risk free rate: 

Assets and Weights

    + SPY (S&P500 fund) weighted 25%
    + EFA (a non-US equities fund) weighted 25%
    + IJS (a small-cap value fund) weighted 20%
    + EEM (an emerging-mkts fund) weighted 20%
    + AGG (a bond fund) weighted 10%
    
Risk Free Rate
    
    + rfr = .0003 or .03%

Let's load our packages: 

```{r setup}
# first install the packages if they are not already installed in your environment.
# install.packages("tidyverse")
# install.packages("tidyquant")
# install.packages("timetk")

library(tidyverse)
library(tidyquant)
library(timetk)

load()
```



```{r}

rfr <- .00008

```


On to the Sortino analysis.  Calculating the Sortino Ratio in the `xts` world is almost depressingly convenient. We call `SortinoRatio(portfolio_returns_xts, MAR = MAR)`, passing our portfolio returns and MAR tot he built-in function from `PerformanceAnalytics`.

```{r}
sortino_xts <- 
  SortinoRatio(portfolio_returns_xts, MAR = MAR) %>% 
  `colnames<-`("ratio")
```

From a substantive perspective, we could stop here and start visualizing with `highcharter`. 

Instead, we will run the calculation by-hand, implementing the equation for the Sortino Ratio via pipes and `dplyr`.  It's not a verbose piped workflow. In short, we call `summarise(ratio = mean(returns - MAR)/sqrt(sum(pmin(returns - MAR, 0)^2)/nrow(.)))`.  

Note the use of `slice(-1)` to remove the first row. I want to delete that first 0.00 for January of 2005 to be consistent with the `xts` operations, but that is an important choice and one that could be questioned. Perhaps we should instead re-wrangle our `xts` object to make it consistent? Either way, we want to be explicit about the choice so that others can reproduce this work later.

```{r}
sortino_byhand <- 
  portfolio_returns_tidy %>% 
  slice(-1) %>%
  summarise(ratio = mean(returns - MAR)/sqrt(sum(pmin(returns - MAR, 0)^2)/nrow(.)))


sortino_byhand

```


Now on to `tidyquant`, which allows us to apply the `SortinoRatio` function from `PerformanceAnalytics` to a `tibble`. As long as we are passing it the same data as we passed originally with the `xts` object, we expect the same result.


```{r}
sortino_tidy <- 
  portfolio_returns_tidy %>%
  slice(-1) %>% 
  tq_performance(Ra = returns, 
                 performance_fun = SortinoRatio, 
                 MAR = MAR,
                 method = "full") %>% 
  `colnames<-`("ratio")
```

Let's compare our 3 Sortino objects. 

```{r}
sortino_xts[1]
sortino_byhand$ratio
sortino_tidy$ratio
```

We have consistent results from `xts`, `tidyquant` and our by-hand piped calculation.  It might feel like a lot of work to get the same result three times but it forced us to look under the hood of the built-in functions and it might serve us well in the future should we have data or a project that fits better with one of the three methods.  

That's all for today. Next time we will visualize the Sortino Ratio and its data slicing implications using `ggplot2` and `highcharter`. Thanks for reading.


