---
title: "Visualizing vol with returns distribution"
output: html_notebook
---

```{r setup, include = FALSE}
library(tidyquant)
library(tidyverse)
library(timetk)
library(tibbletime)
library(highcharter)

knitr::opts_chunk$set(message=FALSE, warning=FALSE)
library(scales)
load("~/reproducible-fin-chapters/book-data.Rdata")
```

We visualize returns with ggplot, and then we visualize skew and kurtosis with ggplot but we skipped volatility. let's do that now by highlighting returns in stand dev and outside standard dev


```{r}
sd <- sd(portfolio_returns_tq_rebalanced_monthly$returns)
mean <- mean(portfolio_returns_tq_rebalanced_monthly$returns)

portfolio_returns_tq_rebalanced_monthly %>%
  mutate(scat_col_red = 
           ifelse(returns < (mean - sd), 
                  returns, NA),
         scat_col_green = 
           ifelse(returns > (mean + sd), 
                  returns, NA),
         scat_col_blue = 
           ifelse(returns > (mean - sd) &
                  returns < (mean + sd),
                  returns, NA)) %>% 
  
  ggplot(aes(x = date)) + 
  
  geom_point(aes(y = scat_col_red),
               color = "red") +
  
  geom_point(aes(y = scat_col_green),
               color = "green") +
  
  geom_point(aes(y = scat_col_blue),
               color = "blue") +
  
  geom_hline(yintercept = (mean + sd), color = "purple", linetype = "dotted") +
  geom_hline(yintercept = (mean-sd), color = "purple", linetype = "dotted") +
  scale_x_date(breaks = pretty_breaks(n = 8)) +

 ylab("percent monthly returns")
```

This is showing us returns over time and whether they fall below or above one standard deviation from the mean.  One observation to me is how many red or green circles we see after 2017. 0! That's zero monthly returns that are least one standard deviation from the mean. When we get to rolling volatility, we should see this reflected as a low rolling volatility through 2017. If we something different, we need to investigate. 

```{r}
portfolio_returns_tq_rebalanced_monthly %>%
  mutate(hist_col_red = 
           ifelse(returns < (mean(returns) - 2*sd(returns)), 
                  returns, NA),
         hist_col_green = 
           ifelse(returns > (mean(returns) + 2*sd(returns)), 
                  returns, NA),
         hist_col_blue = 
           ifelse(returns > (mean(returns) - 2*sd(returns)) &
                  returns < (mean(returns) + 2*sd(returns)),
                  returns, NA)) %>% 
  ggplot() + 
  
  geom_histogram(aes(x = hist_col_red),
               alpha = .7, 
               binwidth = .003, 
               fill = "red", 
               color = "red") +
  
  geom_histogram(aes(x = hist_col_green),
               alpha = .7, 
               binwidth = .003, 
               fill = "green", 
               color = "green") +
  
  geom_histogram(aes(x = hist_col_blue),
               alpha = .7, 
               binwidth = .003, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  
scale_x_continuous(breaks = pretty_breaks(n = 10)) +
xlab("monthly returns")
```


How about a plot to show how different from the riskiest asset, that with the highest standard deviation? 