---
title: "Sortino Notebook"
output:
  pdf_document:
         latex_engine: xelatex
  html_notebook: default
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}

library(tidyquant)
library(tidyverse)
library(timetk)
library(broom)
library(tibbletime)
library(highcharter)
library(scales)

knitr::opts_chunk$set(message=FALSE, warning=FALSE)

load("~/reproducible-fin-chapters/returns/book-data.Rdata")
```



The theory behind the Sortino Ratio is that the riskiness of a portfolio is better measured by the deviation of returns *below* a target return, instead of by the standard deviation of all returns.  This stands in contradistinction to the more used Sharpe Ratio, which measures return/risk by the ratio of the returns above the risk free rate divided by the standard deviation of *all* returns.  By way of history, Harry Markowitz, Nobel laureate and father of MPT whom we mentioned in this sections' introduction, noted that downside deviation might be a better measure of risk than the standard deviation of all returns, but its calculation was computationally too expensive[^1] (it was 1959, if he only he'd had R on his laptop).

[^1]: Markowitz, Harry. Portfolio Selection: Efficient Diversification of Investments, John Wiley & Sons, 1959

The Sortino Ratio equation is as follows:   

$$Sortino~Ratio_{portfolio}=\frac{(\overline{Return_{portfolio}-MAR})}{\sqrt{\sum_{t=1}^n min(R_t-MAR,~0)^2}/n}$$

The denominator in that equation (called the Downside Deviation, semi-deviation or downside risk) can be thought of as the deviation of the returns that fall below some target rate of return for the portfolio. That target rate is called the Minimum Acceptable Rate, or MAR. The numerator is the mean portfolio return minus the MAR. It can be thought of as excess returns in the same was as the numerator of the Sharpe Ratio, except for Sortino it's in excess of whateve minimum rate our team (or our clients) choose.


Let's assign the MAR of .8% to the variable `MAR`. Note that we are holding this portfolio to a higher standard now than we did in the last chapter.

```{r}

MAR <- .008

```

Just as we reused a lot of our skewness code flow in our kurtosis calculation, we will reuse some Sharpe code flow for the Sortino calculations.

In the `xts` world, very similar to the Sharpe Ratio, `PerformanceAnalytics` makes it quick to calculate Sortino. It's the same code as we used for Sharpe, except we call the function `SortinoRatio()` instead of `SharpeRatio()`, and the argument is `MAR = MAR` instead of `Rf = rfr`.

```{r}
sortino_xts <- 
  SortinoRatio(portfolio_returns_xts_rebalanced_monthly, MAR = MAR) %>% 
  `colnames<-`("ratio")
```

We'll again head to `dplyr` and run a by-hand calculation with `summarise(ratio = mean(returns - MAR)/sqrt(sum(pmin(returns - MAR, 0)^2)/nrow(.)))`. This should look familiar from the previous chapter, but with a different equation.


```{r}
sortino_byhand <- 
  portfolio_returns_tq_rebalanced_monthly %>% 
  summarise(ratio = mean(returns - MAR)/sqrt(sum(pmin(returns - MAR, 0)^2)/nrow(.)))
```


Now on to `tidyquant`, which allows us to apply the `SortinoRatio` function from `PerformanceAnalytics` to a `tibble`. 


```{r}
sortino_tidy <- 
  portfolio_returns_tq_rebalanced_monthly %>%
  tq_performance(Ra = returns, 
                 performance_fun = SortinoRatio, 
                 MAR = MAR,
                 method = "full") %>% 
  `colnames<-`("ratio")
```

Let's compare our 3 Sortino objects. 

```{r}
sortino_xts[1]
sortino_byhand$ratio
sortino_tidy$ratio
```

We have consistent results from `xts`, `tidyquant` and our by-hand piped calculation.
Again, the Sortino Ratio is most informative when compared to other Sortino Ratios. Is our portfolio's Sortino good, great, awful? Let's compare it to the Sortino Ratio of the S&P500 in the same time period.

```{r}

```{r, include=FALSE}
market_returns_xts <- 
    getSymbols("SPY", src = 'yahoo', from = "2013-01-01", 
             auto.assign = TRUE, warnings = FALSE) %>% 
    map(~Ad(get(.))) %>% 
    reduce(merge) %>%
    `colnames<-`("SPY") %>% 
    to.monthly(indexAt = "last", OHLC = FALSE) 

market_returns_tidy <-
  market_returns_xts %>% 
    tk_tbl(preserve_index = TRUE, rename_index = "date") %>%
    mutate(returns = (log(SPY) - log(lag(SPY)))) %>% 
    na.omit() %>%
    select(date, returns)
```


```{r}
market_returns_sortino <- 
  market_returns_tidy %>% 
  summarise(ratio = mean(returns - MAR)/sqrt(sum(pmin(returns - MAR, 0)^2)/nrow(.)))

market_returns_sortino$ratio
```

Ok, once again, the S&P500 has beat our portfolio! That makes sense given that the S&P500 also had a higher Sharpe Ratio. 

### Visualizing Sortino

First, we will get a sense for what proportion of our portfolio returns exceeded the MAR.

Very similar to our work on the Shapre Ratio, we will add a column for returns that fall below MAR with `mutate(returns_below_MAR = ifelse(returns < MAR, returns, NA))` and add a column for returns above MAR with `mutate(returns_above_MAR = ifelse(returns > MAR, returns, NA))`. We do this with an eye towards `ggplot`.

```{r}
sortino_byhand <- 
  portfolio_returns_tq_rebalanced_monthly %>% 
  mutate(ratio = mean(returns - MAR)/sqrt(sum(pmin(returns - MAR, 0)^2)/nrow(.))) %>% 
  mutate(returns_below_MAR = ifelse(returns < MAR, returns, NA)) %>%
  mutate(returns_above_MAR = ifelse(returns > MAR, returns, NA))

```

Let's use that new object and it's new columns to create a scatterplot of returns using `ggplot`. 
to again grasp how many of our returns are above the MAR and how many are below the MAR. 

We will create green points for returns above MAR with `geom_point(aes(y = returns_above_MAR), colour = "green")` and red points for returns below MAR with `geom_point(aes(y = returns_below_MAR), colour = "red") `.  

As we did with the Sharpe Ratio, let's add a blue vertical line at November of 2016.  We will also include a horizontal purple dotted line at the MAR.

```{r, warning = FALSE, message = FALSE}
sortino_byhand %>% 
  ggplot(aes(x = date)) +
  geom_point(aes(y = returns_below_MAR), colour = "red") +
  geom_point(aes(y = returns_above_MAR), colour = "green") + 
  geom_vline(xintercept = as.numeric(as.Date("2016-11-30")), color = "blue") +
  geom_hline(yintercept = MAR, color = "purple", linetype = "dotted") +
  annotate(geom="text", x=as.Date("2016-11-30"), 
           y = -.05, label = "Election", fontface = "plain", 
           angle = 90, alpha = .5, vjust =  1.5) +
  ylab("percent monthly returns")
```


It appears that about half of our returns fall below the MAR. Do we consider that to be a successful portfolio? This is not a rigorous test - what strikes us from the number of red dots and where they fall? Do we notice a trend? A period with consistently below or above MAR returns?

Since the election in 2016, there has been only two monthly returns below the MAR and that will lead to a large Sortino since November. 

Next we will build a histogram of the distribution of returns with `geom_histogram(alpha = 0.25, binwidth = .01, fill = "cornflowerblue")`. We will again add a line for the MAR.

```{r}
sortino_byhand %>% 
  ggplot(aes(x = returns)) +
  geom_histogram(alpha = 0.25, binwidth = .01, fill = "cornflowerblue") +
  geom_vline(xintercept = MAR, color = "green") +
  annotate(geom = "text", x = MAR,
    y = 22, label = "MAR", fontface = "plain", angle = 90, alpha = .5, vjust =  1)

```

The Sortino Ratio is often accompanied by a density plot and we'll build one now. First, we will start simple with `stat_density(geom = "line", size = 1, color = "cornflowerblue")` to create a `ggplot` object called `sortino_density_plot`.

```{r}

sortino_density_plot <- 
  sortino_byhand %>% 
  ggplot(aes(x = returns)) +
  stat_density(geom = "line", size = 1, color = "cornflowerblue") 

sortino_density_plot
```

A slight negative skew is a bit more evident here. It would be nice to shade the area that falls below the MAR.  To do that, let's create an object called `shaded_area` using `ggplot_build(p)$data[[1]] %>% filter(x < MAR)`. That snippet will take our original `ggplot` object and create a new object filtered for x values less than MAR. Then we use `geom_area` to add the shaded area to `sortino_density_plot`.

```{r}

shaded_area_data <- ggplot_build(sortino_density_plot)$data[[1]] %>% 
  filter(x < MAR)

sortino_density_plot_shaded <- sortino_density_plot + 
  geom_area(data = shaded_area_data, aes(x = x, y = y), fill="pink", alpha = 0.5) 

sortino_density_plot_shaded
```


Let's add a vertical line label at the exact MAR and an arrow to tell people where downside volatility resides. Note how we can keep adding layers to the `sortino_density_plot_shaded` object from above, which is one of great features of `ggplot`. It allows experimentation with aesthetics without changing the core plot with each iteration.

```{r}

sortino_density_plot_shaded +
  geom_segment(aes(x = 0, y = 1.5, xend = -.05, yend = 1.5),
  arrow = arrow(length = unit(0.4, "cm")), size = .05) +
  geom_segment(data = shaded_area_data, aes(x = MAR, y = 0, xend = MAR, yend = y), 
               color = "red", linetype = "dotted") +
  annotate(geom = "text", x = MAR, y = 5, label = "MAR = 0.8%", 
           fontface = "plain", angle = 90, alpha = .8, vjust =  -1) +
  annotate(geom = "text", x = -.02, y = .1, label = "Downside", 
           fontface = "plain", alpha = .8, vjust =  -1)
```

As with our scatterplot, we have not been shy about aesthetic layering but one goal here is to explore `ggplot` tools, which gives us license to be overinclusive.  

We have done some good work for visualizing the portfolio's returns and how they are distributed relative to the MAR, and how the MAR separates part of the the returns to downside risk. That gives us some intution about the Sortino Ratio.

As with Sharpe Ratio, let's get to some rolling Sortinos. There is a slight wrinkle though - remember that we exclude above MAR returns from the denominator. If our rolling window is too small, we might end up with a denominator that includes 1 or 2 or 0 downside deviations. That would accurately reflect that the portfolio has done well in that small window but it might report a misleadingly huge number for the rolling window. The rolling 6 month demonstrates this. 

First, we need to calculate the rolling 6-month Sortino with `rollapply(portfolio_returns_xts, 6, function(x) SortinoRatio(x))`. Then we can visualize with `highcharter`.

```{r}

portfolio_sortino_roll_6_xts <- 
  rollapply(portfolio_returns_xts_rebalanced_monthly, 6, 
            function(x) SortinoRatio(x, MAR = MAR)) %>% 
  `colnames<-`("6-rolling")


```

Let's take a quick peek at the rolling 6-month Sortino visualization and investigate other unusual occurrences when we slice the data.

```{r}
highchart(type = "stock") %>%
  hc_title(text = "Rolling Sortino") %>%
  hc_add_series(portfolio_sortino_roll_6_xts, name = "Sortino", color = "cornflowerblue") %>%
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE)
```

The rolling 6-month has so many bizarre spikes. It nicely emphasizes how we need to be careful with the Sortino Ratio, short time periods and rolling applications.

Let's see how the rolling 24-month compares.

```{r}

portfolio_sortino_roll_24_xts <- rollapply(portfolio_returns_xts_rebalanced_monthly, 24, 
                           function(x) SortinoRatio(x, MAR = MAR))

highchart(type = "stock") %>%
  hc_title(text = "Rolling Sortino") %>%
  hc_add_series(portfolio_sortino_roll_24_xts, name = "Sortino 24", color = "lightgreen") %>%
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE)
```

Ah, much better. We can see the movements and how the Sortino has changed through the life of this portfolio, but within a reasonable range of .6 to -0.4.

Let's compare to the rolling S&P500 Sortino. 

```{r}
market_sortino_roll_24_xts <- rollapply(market_returns_xts, 24, 
                           function(x) SortinoRatio(x, MAR = MAR))
  

highchart(type = "stock") %>%
  hc_title(text = "Rolling sharpe") %>%
  hc_add_series(portfolio_sortino_roll_24_xts, name = "Portfolio", color = "cornflowerblue") %>%
  hc_add_series(market_sortino_roll_24_xts, name = "SP500", color = "green") %>%
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE)

```


Rolling Sortinos need to be handled with care but there are a few nice payoffs. First, these charts force us and our end users to reflect on how time periods can affect Sortino to extremes. Be skeptical if someone reports a fantastic 6-month Sortino. Second, as an exploratory device, the rolling ratios highligh time periods deserving of more investigation.  Third, with Sortino (and Sharpe) Ratios, there's a temptation to look at the final number for a portfolio's life and judge it 'good' or 'bad'. These rolling visualizations can help reframe the analysis and look at how the portfolio behaved in different economic and market regimes.


### On to Shiny

First off, have a look at the Shiny App and notice how similar it is to the Sharpe Ratio app in structure. We just saw how the calculations are different, but from a Shiny perspective, the only difference is that the end user chooses a MAR instead of a risk-free rate. 

```{r, eval = FALSE}

fluidRow(
  column(5,
  numericInput("mar", "MAR%", .8, min = 0, max = 3, step = .01)),
  column(5,
  numericInput("window", "Window", 6, min = 2, max = 24, step = 2))
)
```

Note that we ask for a MAR *percent*. Thus, when calculating the reactive, we divide by 100. 

```{r, eval = FALSE}
mar <- eventReactive(input$go, {input$mar/100})
```

Why do that? No substantive reason - perhaps it's more intuitive to the user.

On that same theme, I removed the value boxes so the user doesn't get to see the overall Sortino Ratio for the portfolio and the market. Why did I do that? 

Let's talk testing and the user experience here. 